---
title: "Covid91 vaccine study Final2024"
author: "Jovygomez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


# Males

## Methods

Question: Is the Covid-91 effective in males? 

$H_0$:  The vaccine is not effective in preventing covid-91 in males.

$H_A$:  The vaccine is effective in preventing covid-91 in males 

```{r}
Male <- FinalDat[FinalDat$sex=='M',]
```


```{r}
barchartGC(~infected + treatment,data=Male)
barchartGC(~infected  + treatment,data=Male, type="percent")

```

From the bar chart it shows that you are less likely to contract covid-91 if you did receive the vaccine.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```
This table shows percentages of people who contracted COVID-91 and those who did not (OK) for both the placebo and the vaccine groups.  

Placebo Group: 73.80% of individuals on the placebo got infected with COVID-91, while 50.01% remained healthy (OK).
Vaccine Group: Only 26.20% of individuals who received the vaccine got infected, while 49.99% remained healthy  

In the placebo group, a higher percentage of participants (73.80%) contracted COVID-91 compared to the vaccine group (26.20%), indicating the vaccine significantly reduced the likelihood of infection. 
The total population breakdown further confirms that the incidence of COVID-19 was significantly lower in the vaccine group (1.02%) compared to the placebo group (2.82%).


## Infrential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

This data and the Fisher exact test shows that the P-Value is 2.2e-16 which is
significantly lower than 0.05 which means we reject the null hypothesis.


## Conclusion

The bar chart shows that you are less likely to contract covid91 if you did
receive the vaccine. The numerical data shows the results in percentages. Within
the placebo group 73.80% of individuals got infected with covid91. While in the
vaccine group 26.20% contracted covid-91. This indicates that receiving the
vaccine as a male significantly reduces the risk of contracted covid-91. The
inferential data including the Fisher exact test shows that the P-Value is
2.2e-16 which is significantly lower than 0.05 which means we reject the null
hypothesis. All in all these examples and simulations show that is as a male you
are far less likely to contract covid-91 upon receiving the vaccine.

# Females


## Methods. 

Question: Is the Covid-91 effective in females? 


$H_0$:  The vaccine is not effective in preventing covid 91 in females? 

$H_A$:  The vaccine is effective in preventing covid 91 in females?  

infected~sex(F)
categorical~categorical

Well be using Cat~Cat 

## Graphical Results 
```{r}
Female <- FinalDat[FinalDat$sex=='F',]
```


```{r}
barchartGC(~infected + treatment,data=Female)
barchartGC(~infected + treatment,data=Female, type="percent")

```

This bar chart shows that females are less likely to contract Covid91 after
reviewing the vaccine. 


## Numerical Results 

```{r}
table1 <- xtabs(~infected + treatment, data=Female)
rowPerc(table1)
colPerc(table1)
```

The numerical data presents the results in percentages. In the placebo group, 64.93% of individuals contracted COVID-91, while in the vaccine group 118, only 35.07% were infected. This suggests that receiving the vaccine substantially reduces the risk of contracting COVID-91, particularly for females.

## Infrfential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

This Chi-Square test shows a P-Value of 1.914e-08 and the Fisher exact test shows that the P-Value is 1.246e-08 which is significantly lower than 0.05 which means we reject the null hypothesis. 


## Conclusion

The bar chart shows that you are less likely to contract Covid91 if you did
receive the vaccine. The numerical data shows the results in percentages. 
Within the placebo group 64.93% of individuals got infected with Covid91, while in the
vaccine group 35.07% contracted covid-91.This indicates that receiving the vaccine as a female significantly reduces the risk of contracted Covid-91. The inferential data including the Fisher exact test shows that the P-Value is 1.246e-08 which, is significantly lower than 0.05 which means we reject the null hypothesis In summation these examples and simulations show that is as a female you are far less likely to contract Covid-91 if you receive the vaccine.


# LGBTQ

## Methods. 

Question: Is the Covid-91 vaccine effective in preventing Covid91 in members of the LGBTQ community? 


$H_0$:  The vaccine is not effective in preventing covid-91 in members of the LGBTQ community.

$H_A$:  The vaccine is effective in preventing covid 91 in members of the LGBTQ community.  

infected~LGBT(gay)
categorical~categorical

Well be using Cat~Cat


## Graphical Results

```{r}
LGBTQ <- FinalDat[FinalDat$LGBTQ== 'gay',]
```



```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

This bar chart shows that members of the LGBTQ community are less likely to contract Covid91 after reviewing the vaccine.


## Numerical Results

```{r}
table1 <-xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```


The numerical data shows the results in percentages. Within the placebo group 80.00% of individuals got infected with Covid-91. While in the vaccine group 20.00% contracted Covid-91. This indicates that receving the vaccine as a LGBTQ community member significantly reduces the risk of contracted Covid-91.


## Infrential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi-square test shows hat the P-value is 2.266e-06 and data the Fisher exact test shows that the P-value is 7.269e-07 which is significantly lower than 0.05 which means we reject the null hypothesis. 

## Conclusion

The bar chart shows you are less likely to contract Covid-91 if you did recieve the vaccine. The numerical data shows the results in percentages. Within the placebo group 80.00% of individuals got infected with Covid-91. While in the vaccine group 20.00% contracted Covid-91. This indicates that recieving the vaccine as a LGBTQ member significantly reduces the risk of contracted Covid-91. The infrential data shows Chi-squared P-value of 2.266e-06, this data also includes the Fisher exact test that shows that the P-Value is 7.269e-07 which is significantly lower than 0.05 which means we reject the null hypothesis. In conclusion these examples are simulations show that a member of the LGBTQ community you are far less likely to contract Covid-91 if you recieve the vaccine.  

# Druggies


## Methods. 

Questions: Is the Covid-91 effecitve in Druggies? 

$H_0$:  The vaccine is not effective in preventing covid-91 in a drug user? 

$H_A$:  The vaccine is effective in preventing covid-91 in a drug user?  

infected~DrugUser(yes)
categorical~categorical

Well be using Cat~Cat

## Graphical Results 


```{r}
DrugUser <-FinalDat[FinalDat$DrugUser=='yes' ,]
```

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")

```

According to the bar chart you are more likely to get Covid91 if reciece the vaccine as a drug user.

##Numerical Results 

```{r}
table1 <-xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows the results in percentages. Within the placebo group 8.33% of individuals got infected with covid91. While in the vaccine group 91.67% contracted covid91. This indicates that recieving the vaccine as a drug user significantly increases the risk of contracted covid91. 

##Infrential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The Chi-Square test shows a P-Value of 1.414e-10 data and the Fisher exact test shows that the P-value is 3.261e-12 which is significantly lower than 0.05 which means we reject the null hypothesis. 

##Conclusion

The bar chart 

# Overall Results and Conclusions